---
title: "HW9_504"
author: "Xuchen Wang"
date: "March 23, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

#### Problem 3 (a)
```{r}
data <- read.table("/Users/xuchenwang/nba_data.txt", header = T)
nba <- data.frame(data$height,data$weight)
x <- unname(as.matrix(nba))
# center the data
avg <- colMeans(x)
x <- t(t(x)-avg)
# plot the data
plot(x,xlab = "height of the player", ylab = "weight of the player")
eigens <- eigen(t(x)%*%x)
vec1 <- eigens$vectors[,1]
vec2 <- eigens$vectors[,2]
abline(a=0, b=vec1[2]/vec1[1],col="red")
abline(a=0, b=vec2[2]/vec2[1],col="blue")
```

#### (b)
```{r}
proj_dom <- as.vector(vec1%*%t(x))
extre1 <- order(proj_dom)[c(1,nrow(x))]
data$player[extre1]
```
John Lucas is tall and fat. Nikola Pekovic is short and thin.


```{r}
proj_sec <-  as.vector(vec2%*%t(x))
extre2 <- order(proj_sec)[c(1,nrow(x))]
data$player[extre2]
```



#### (c)
```{r}
mean(proj_dom^2)
eigens$values[1]/nrow(x)
```

#### (d)
```{r}
z <- cbind(proj_dom,proj_sec)
t(z)%*%z
eigens$values
```

#### Problem 4 (a)
```{r}
data2 <- read.table("/Users/xuchenwang/votes_formatted.txt", header = T)
votes <- unname(data2[,-1])
data3 <- read.table("/Users/xuchenwang/senators_formatted.txt", header = T)
party <- data3$party
# center the data
avg <- colMeans(votes)
votes <- t(t(votes)-avg)
x <- votes%*%t(votes)

power_iter <- function(A, eps= 10^(-10), max_i= 10^4){
  # This function use power iteration to calculate the dominant eigenvector
  # parameter: A (original matrix)
  #            eps (error of the stop condition)
  #            max_i (maximum iterations)
  # return: v (dominant eigenvector)
  #         i (number of iterations)
  n <- ncol(A)                          # number of pages
  v <- rep(1,n)                         # randomly generate an initial vector
  i <- 0
  sign <- TRUE                          # use sign to control the process of loop
  while(sign){
    u <- v
    v <- A%*%v                               
    v <- v/norm(v)                      # apply equation and normalize 
    i <- i+1
    if (norm(v-u)< eps || i >= max_i)
      {sign <- FALSE}                   # meet stop condition, then jump out of loop
    }
  return(list(domi_vector=v, iterations=i))
}
dom_vec <- as.vector(power_iter(x)$domi_vector)
proj_dom <- as.vector(t(votes)%*%dom_vec)

proj_1d <- data.frame(proj_dom, party)
ggplot(data = proj_1d)+
  geom_point(aes(x=1:100,y=proj_dom,color=factor(party)))+
  xlab("index of senators")+
  ylab("Dominant eigenvector")
```

From the graph, it is imformative since it could roughly separate different parties.

#### (b)
```{r}
power_iters <- function(A, k, eps = 10^(-10), max_i = 10^4){
  # This function use power iteration to calculate the first k dominant eigenvector
  # parameter: A (matrix)
  #            k (number of dominant eigenvectors)
  #            eps (error of the stop condition)
  #            max_i (maximum iterations)
  # return: v (dominant eigenvector)
  #         i (number of iterations)
  n <- ncol(A)
  set.seed(1)
  Q <- matrix(rnorm(k*n),ncol = k, nrow = n)
  i <- 0
  sign <- TRUE
  while(sign){
    U <- Q
    M <- A%*%Q
    Q <- qr.Q(qr(M))
    i <- i+1
    if (max(abs(Q-U))< eps || i >= max_i)
    {sign <- FALSE}  
    }
  return(list(dom_vecs=Q, iterations=i))
}
Q <- power_iters(x,k=2)$dom_vecs
proj <- t(votes)%*%Q

proj_2d <- data.frame(proj[,1],proj[,2],party)
ggplot(data = proj_2d)+
  geom_point(aes(x=proj[,1],y=proj[,2],color=factor(party)))+
  xlab("Dominant eigenvector")+
  ylab("Second dominant eigenvector")
```

#### (c)
```{r}
power_iters(x,k=10)
eigens <- eigen(x)
eigens$vector[,1:10]
plot(eigens$values)
plot(eigens$values[1:10],xlab = "index of lambda", ylab = "eigen values")
# we can compute all eigenvectors using power iteration
eigenvec <- power_iters(x,k=10)
```

There is an obvious dimensional reduction that would capture a significant amount of variance. Because the eigenvalues drop quickly at lambda1 to lambda3 and remains approximately constant from lambda4.







