---
title: "hw11_504"
author: "Xuchen Wang"
date: "April 5, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Problem 2
```{r}
options(digits = 16)
i <- -20:0
h <- 10^i
d1 <- (exp(0+h)-exp(0))/h
d2 <- (exp(0+h)-exp(0-h))/(2*h)

correct_digits <- function(x,obj){
  error <- abs(x-obj)
  for(i in 1:length(error)){
    if(error[i] >= 0.1){j <- 0}
    else{
      j <- 1
      while(error[i]<10^(-j)){
        j <- j+1
        if(j>16){break}
        }
      j <- j-1
      }
    print(paste(i-21,':','The error is',error[i],'and correct digit is',j))
    }
}
# show the finite difference and their correct digits
d1
correct_digits(d1,1)
d2
correct_digits(d2,1)
```

Explain the results:
For the two point finite difference, the optimal h is 10^-8 and the corresponding error is about 6*10^-9, the correct digit is 8, which equals the optimal h by calculating the error theoretically. For the two-side estimation, the optimal h is 10^-5, which is nearly 10^(-16/5), the corresponding error is about 10^-11, which is nearly 10^(-32/3).
Thus, the results of both estimation verify the optimal h and error by calculating the error in class.

#### Problem 4
```{r}
Fapprox <- function(n,method,cutoff=3){
  f <- function(x){1/sqrt(2*pi)*exp(-x^2/2)}
  h <- cutoff/n
  x1 <- (0:(n-1))*h
  x2 <- (1:n)*h
  if(method=="reimann"){return(sum(f(x1)*h))}
  else if(method=="trapezoid"){return(sum((f(x1)+f(x2))*h/2))}
  else if(method=="useR"){return(integrate(f,lower = 0,upper = cutoff,subdivisions = n))}
}

# using reimann method
Fapprox(10,method = "reimann")
Fapprox(100,method = "reimann")
Fapprox(1000,method = "reimann")
Fapprox(10000,method = "reimann")
# using trapezoid method
Fapprox(10,method = "trapezoid")
Fapprox(100,method = "trapezoid")
Fapprox(1000,method = "trapezoid")
Fapprox(10000,method = "trapezoid")
# using R method
Fapprox(10,method = "useR")
Fapprox(100,method = "useR")
Fapprox(1000,method = "useR")
Fapprox(10000,method = "useR")

```

Compare accuracy:
For the Reimann method, the accuracy improves obviously as n increases. For the Trapezoid method, the accuracy improves little as n increases and almost remains the same when n is larger than 1000. For the R method, the accuracy remains the same. Thus, R method is the most accurate method. When n is small, we use R method to estimate. The Trapezoid is the second accurate method, when n is sufficient large, such as 1000, we can use both R method and Trapezoid method to estimate. When n is very very large, we can use all of those three methods to estimate.














